---
title: "Most Damaging Severe Weather Across USA"
output: html_document
---

#Synopsis
By analysing the NOAA Storm database, and cleanning up the data extensively by removing inconsistencies, we have reached the following conclusions:

  1-The most harmful severe weather to the human health is tornados
  
  2-The most damaging severe weather to human properties and to the crops is floods.

#Preprocessing Data
##Loading required libraries
```{r libraries loading}
library(ggplot2)
library(graphics)
library(xtable)
```

##Downloading the data
```{r data downloading}
fileUrl = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(fileUrl, destfile = "..//data//repdata_data_StormData.csv.bz2", method = "auto")
file.bz <- bzfile("..//data//repdata_data_StormData.csv.bz2","rt")  

stormDataDownloaded = date()
```

##Loading the data

```{r data loading}
#stormData = read.csv("..//data//repdata_data_StormData.csv", sep = ",")

stormData = read.csv(file.bz)
```

##Summary of Storm Data
```{r data summary}
summary(stormData)
```

#Data Processing
```{r data processing}
# Taking the sum of fatalities for each event type
fatalities = tapply(stormData$FATALITIES, stormData$EVTYPE, sum)
# Taking the sum of injuries for each event type
injuries = tapply(stormData$INJURIES, stormData$EVTYPE, sum)
# Taking the sum of property damage for each event type
propertyDamage = tapply(stormData$PROPDMG, stormData$EVTYPE, sum)
# Taking the sum of crop damage for each event type
cropDamage = tapply(stormData$CROPDMG, stormData$EVTYPE, sum)

#creating a new dataframe to store all economic and health damages for each type
consequences = NULL
consequences  = data.frame(event = dimnames(fatalities), injuries, fatalities, propertyDamage, cropDamage)

names(consequences) = c("event", "injuries", "fatalities", "propertyDamage", "cropDamage")

# A simple summation of all fatalities and injuries can be a measure of how harmful the event is
consequences$health = consequences$injuries + consequences$fatalities

healthPlot = ggplot(consequences, aes(x=event, y=health)) + geom_bar(stat="identity")

healthPlot += 

# A simple summation of all property and crop damages can be a measure of how damaging the event is
consequences$economic = propertyDamage + cropDamage
```

```{r}
sortedHealth=consequences[order(consequences$health, decreasing = TRUE),c(1,2,3,6)]
sortedEconomic=consequences[order(consequences$economic, decreasing = TRUE),c(1,4,5,7)]
#sortedEconomic = sort()
```

#Results
##The types of events that are most harmful to population health
`r consequences[which(consequences$health == max(consequences$health)), 1]`
###Its harm can lead to a max of:
`r max(consequences$health) ` injuries and fatalities

##The types of events that have the greatest economic consequences
`r consequences[which(consequences$economic == max(consequences$economic)), 1] `
###Its damage is maxed at: `r max(consequences$economic) `

```{r table}
#print(xtable(consequences), type="latex")

```


```{r figure, fig.width= 480, fig.height= 480, echo=FALSE}

```


