---
title: "Severe Weather Events Report"
output: pdf_document
---



##Loading required libraries
```{r}
library(ggplot2)
```

## Download the data
```{r}
#fileUrl = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
#download.file(fileUrl, destfile = "..//data//StormData.zip", method = "auto")
#stordDataDownloaded = date()
```

## Loading the data

```{r}
stormData = read.csv("..//data//repdata_data_StormData.csv", sep = ",")
```

###Summary of Storm Data
```{r}
str(stormData)
```

## Preprocessing the data
```{r}
fatalities = tapply(stormData$FATALITIES, stormData$EVTYPE, sum)
injuries = tapply(stormData$INJURIES, stormData$EVTYPE, sum)
propertyDamage = tapply(stormData$PROPDMG, stormData$EVTYPE, sum)
cropDamage = tapply(stormData$CROPDMG, stormData$EVTYPE, sum)

consequences  = data.frame(dimnames(fatalities), injuries, fatalities, propertyDamage)

consequences$health = consequences$injuries + consequences$fatalities
consequences$economic = propertyDamage + cropDamage
```

The most harmful weather with respect to population health:
`r consequences[which(consequences$health == max(consequences$health)), 1]`
Its harm can lead to a max of:
`r max(consequences$health) ` injuries and fatalities

The most damaging weather events to the economy: `r consequences[which(consequences$economic == max(consequences$economic)), 1] `
Its damage is maxed at: `r max(consequences$economic) `



```{r, echo=FALSE}

```


